---
title: "Defining genetic population clusters using the 1240k sites"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Defining genetic population clusters

For our analysis we need to group individuals for e.g. get frequencies of Neandertal ancestry. We can do this by time. For this we rely on radiocarbon dates. But we can also do it for populations. To define population on a continental or sub-continental level one could simply take the sampling location. But since people tend to move around during (pre-) history we know that people at the same sample location over time are not necessarily genetically close. Hence we are grouping individuals using their genetic information. To do that efficiently we use sites of the 1240k array which is especially good in capturing in-human diversity (but bad for Archaic ancestry). We use the genotype likelihoods provided by admixfrog for ancient individuals (here we take the genotype with the highest LH) and the genotypes given for the SGDP individuals, to calculate all pairwise outgroup F3 statistics with MButi as the outgroup. F3's for all SGDP individuals are done on their pre-assigned SGDP populations.

The Meta_Data_individuals.csv which you can get from the Dryad repo already has the clustered populations!!!

## Functions

```{r eval=T,echo=F, result='hide'}
suppressPackageStartupMessages({
library(reshape2) 
library(dendextend)
library(paletteer)
library(tidyverse)
library(viridis)
library(ggrepel)})


############## change path  ##################
folder_path = "Result_folder_get_pops"

path_to_F3_input_file = "Results_F3_A1240k"

path_to_dryad_downloaded_files = "Dryad_folder"

##############################################

'%notin%' <- Negate('%in%')

HC_Dendrogram <- function(F3_matrix_dis,max_tree_hight=NULL,k=NULL,only_cluster=F){
  hc <- hclust(as.dist(F3_matrix_dis))
  dend_prunned <- as.dendrogram(hc)
  
  
  cut_tree_at = max_tree_hight
  k_clust = k
  
  if((is.null(cut_tree_at) && is.null(k_clust))){
    avg_col_dend <- as.dendrogram(hc) %>% 
      dendextend::set("branches_lwd", 3) 
    
  }
  if(is.null(k_clust) && !is.null(cut_tree_at)) {
    avg_col_dend <- as.dendrogram(hc) %>% dendextend::color_branches(h = cut_tree_at,) %>%
      dendextend::set("branches_lwd", 3) %>% dendextend::color_labels(h = cut_tree_at)
  }
  if(is.null(cut_tree_at) && !is.null(k_clust)) {
    avg_col_dend <- as.dendrogram(hc) %>% dendextend::color_branches(k = k_clust) %>%
      dendextend::set("branches_lwd", 3) %>% dendextend::color_labels(k = k_clust)
  }
  if(only_cluster == T && !is.null(cut_tree_at)){
    cluster_groups = data.frame(group = cutree(hc, k = NULL, h = cut_tree_at),pop = names(cutree(hc, k = NULL, h = cut_tree_at)))
    return(cluster_groups)
  }
  if(only_cluster == T && !is.null(k_clust)){
    cluster_groups = data.frame(group = cutree(hc, k = k_clust, h = NULL),pop = names(cutree(hc, k = k_clust, h = NULL)))
    return(cluster_groups)
  }else{
    plot(avg_col_dend)
    abline(h=cut_tree_at)
  }
}

reorder_F3 <- function(F3_dis){
  dd <- as.dist(F3_dis)
  hc <- hclust(dd)
  F3_dis <-F3_dis[hc$order, hc$order]
}


```

## Read in Data

The calculations for all pairwise F3 are done using ADMIXTOOLS2 (in the script Calculate_F3_on_1240k.Rmd). We only consider sites where at least half of the ancient individuals have SNP information. The F2 from which all the F3's are calculated are done so with parameters:
 * maxmiss = 1
 * auto_only = T
 * adjust_pseudohaploid = FALSE
 * fst=FALSE
 * apply_corr=FALSE
 
We construct a n x n matrix of pairwise F3's distances. The distance is given by $max(F_3) - F_3$, the diagonal is put to zero. We remove genetically closely related individuals (here we only keep the son1 of the Afanasievo family).

```{r eval=T,echo=F, result='hide'}
# sample meta information
Joint_Meta <- read.csv(paste0(path_to_dryad_downloaded_files,"/Meta_Data_individuals.csv",sep = ",")) %>% mutate(F3_pop_annotation = ifelse(ML_BP_Mean == 0 ,SGDP_pop,sample_name)) 
Joint_Meta = Joint_Meta %>%  mutate(ID = c(Joint_Meta$pop[Joint_Meta$ML_BP_Mean == 0],Joint_Meta$sample_name[Joint_Meta$ML_BP_Mean > 0]))

# All together

res_f3_gtLH_Autosomes <- read.csv(paste0(path_to_F3_input_file,"/res_f3_AF_gtLH_A1240k_Shared_Map.csv"))
res_f3_gtLH_Autosomes[res_f3_gtLH_Autosomes == "Salkhit"] <- "SalkhitArchAdm"
res_f3_gtLH_Autosomes_matrix_full = res_f3_gtLH_Autosomes %>% dplyr::select(c(pop2,pop3,est)) %>% pivot_wider(names_from = pop3, values_from = est) %>% column_to_rownames(.,"pop2") %>% as.matrix()
res_f3_gtLH_Autosomes_red <-  res_f3_gtLH_Autosomes %>% filter(pop2 %notin% c("AfanasievoSon2","AfanasievoFather","AfanasievoMother"),pop3 %notin% c("AfanasievoSon2","AfanasievoFather","AfanasievoMother"))
res_f3_gtLH_Autosomes_matrix_red = res_f3_gtLH_Autosomes_red %>% dplyr::select(c(pop2,pop3,est)) %>% pivot_wider(names_from = pop3, values_from = est) %>% column_to_rownames(.,"pop2") %>% as.matrix()

res_f3_gtLH_Autosomes_matrix_full_sim <- res_f3_gtLH_Autosomes_matrix_full
diag(res_f3_gtLH_Autosomes_matrix_full_sim) <- 0
res_f3_gtLH_Autosomes_matrix_full_dis <- max(res_f3_gtLH_Autosomes_matrix_full_sim) - res_f3_gtLH_Autosomes_matrix_full_sim
diag(res_f3_gtLH_Autosomes_matrix_full_dis) <- 0

res_f3_gtLH_Autosomes_matrix_red_sim <- res_f3_gtLH_Autosomes_matrix_red
diag(res_f3_gtLH_Autosomes_matrix_red_sim) <- 0
res_f3_gtLH_Autosomes_matrix_red_dis <- max(res_f3_gtLH_Autosomes_matrix_red_sim)- res_f3_gtLH_Autosomes_matrix_red_sim
diag(res_f3_gtLH_Autosomes_matrix_red_dis) <- 0
```
 
## Heatmap

We can visualize the resulting matrix in a heatmap

```{r eval=T,echo=F, result='hide'}
F3mat <- reorder_F3(res_f3_gtLH_Autosomes_matrix_full_dis)
melted_F3_matrix <- melt(F3mat)
melted_F3_matrix_anno <- melted_F3_matrix %>% inner_join(.,Joint_Meta[,c("ID","SGDP_Superpop","ML_BP_Mean")],by=c("Var1"="ID")) %>% dplyr::rename(Var1_SGDP_Superpop = SGDP_Superpop,Var1_ML_BP_Mean = ML_BP_Mean) %>%
  inner_join(.,Joint_Meta[,c("ID","SGDP_Superpop","ML_BP_Mean")],by=c("Var2"="ID")) %>% dplyr::rename(Var2_SGDP_Superpop = SGDP_Superpop,Var2_ML_BP_Mean = ML_BP_Mean) %>% distinct() %>%
  filter(Var1_SGDP_Superpop != "Africa" & Var2_SGDP_Superpop != "Africa")
melted_F3_matrix_anno = melted_F3_matrix_anno[ order(match(melted_F3_matrix_anno$Var1, melted_F3_matrix$Var1)), ]
melted_F3_matrix_anno$Var1 <- ordered(melted_F3_matrix_anno$Var1, levels = levels(melted_F3_matrix$Var1))
melted_F3_matrix_anno = melted_F3_matrix_anno[ order(match(melted_F3_matrix_anno$Var2, melted_F3_matrix$Var2)), ]
melted_F3_matrix_anno$Var2 <- ordered(melted_F3_matrix_anno$Var2, levels = levels(melted_F3_matrix$Var2))
melted_F3_matrix_anno$value <- ifelse(melted_F3_matrix_anno$Var1 == melted_F3_matrix_anno$Var2 , NA, melted_F3_matrix_anno$value)
colnames(melted_F3_matrix_anno)[colnames(melted_F3_matrix_anno) == "value"] <- "max(F3) - F3"


ggHeatmap <- ggplot(data = melted_F3_matrix_anno, aes(x=Var1, y=Var2, fill=`max(F3) - F3`)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = mean(melted_F3_matrix_anno$`max(F3) - F3`,na.rm = T), limit = c(min(melted_F3_matrix_anno$`max(F3) - F3`,na.rm = T),max(melted_F3_matrix_anno$`max(F3) - F3`,na.rm = T)), space = "Lab")+
  geom_point(data= melted_F3_matrix_anno, aes(x=Var1, y=Var2),alpha=0) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 8, hjust = 1),
        axis.text.y = element_text( vjust = 1, 
                                    size = 8, hjust = 1)) +
  xlab("") + 
  ylab("") +
  coord_fixed()

ggHeatmap

ggsave(filename = paste(folder_path,"/Defining_genetic_populations_F3_dist_ggHeatmap_using_1240k.png",sep=""),ggHeatmap,device = "png",width = 14,height =  14)
```


## Clustering

We are using hierarchical clustering to form groups out of the F3 distances between individuals.

### Predefin clusters

From literature we know that there are already some either very distinct clusters, individuals that did not contribute to later times or that contribute equally to most non-African populations today. We assign these individuals predefined clusters. 

 * Supercluster = Oceania; cluster = Oceania: Papuans, Bougainville, Australia and Leang Panninge
 * Supercluster = Africa ; cluster = Africa: all sub-Saharan Africans
 * Supercluster = Africa; cluster = North-Africa: Saharawi, Mozabite
 * Supercluster = EarlyOoA, cluster = EarlyOoA: Oase, ZlatyKun, BKBB7240, BKCC7335, BKF6620, UstIshim

We exclude these predefined clusters form our hierarchical clustering

```{r eval=T,echo=F, result='hide'}
Oceania_cluster <- data.frame(samples = c("Papuan","Australian","Bougainville","LeangPanninge"),Supercluster= "Oceania",cluster= "Oceania")

Not_clustered <- data.frame(samples = c("OaseNew","ZlatyKunShotgun","BKBB7240","BKCC7335","BKF6620","UstIshim"),Supercluster= "EarlyOoA",cluster= "EarlyOoA")

Africa <- data.frame(samples = c("Saharawi","Mozabite"),Supercluster= "Africa",cluster= "North_Africa")

exclude_sample <- rownames(res_f3_gtLH_Autosomes_matrix_red)[rownames(res_f3_gtLH_Autosomes_matrix_red_dis) %notin% c(Oceania_cluster$samples,Not_clustered$samples,Africa$samples)]
```

### Superclusters

First we define two superclusters as continental clusters from the remaining individuals.
The resulting cluster are:
 * West-Eurasia and Central Asia
 * Asia and America

```{r eval=T,echo=F, result='hide'}
HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[exclude_sample,exclude_sample],k = 2)

hc_Supercluster_ALL <- HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[exclude_sample,exclude_sample],k=2,only_cluster = T)
colnames(hc_Supercluster_ALL) <- c("Supercluster","samples")
hc_Supercluster_ALL$Supercluster[hc_Supercluster_ALL$Supercluster == hc_Supercluster_ALL$Supercluster[hc_Supercluster_ALL$samples == "French"]] <- "WEurCAS"
hc_Supercluster_ALL$Supercluster[hc_Supercluster_ALL$Supercluster == hc_Supercluster_ALL$Supercluster[hc_Supercluster_ALL$samples == "Han"]] <- "Americas&Asia"


```

### Clusters

Then we define 9 more population clusters. We check for consistency with previous studies. The resulting clusters are broad since there is no point in clusters containing only one or two individuals for frequency estimats.
The clusters are:
 * Siberia&Americas
 * EAS
 * WEur
 * SA
 * AncientEAS
 * Satsurblia
 * ANS (Ancient north Siberians)
 * ANS (Ancient north Eurasians)
 * postLGM-WEurHG (post LGM west Eurasian Hunter Gatherers)
 * preLGM-WEurHG (pre LGM west Eurasian Hunter Gatherers)
 * SWA (South-West Asians)
 
The Afanasievo family is classified as Yamnaya.
 

```{r eval=T,echo=F, result='hide'}
HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[exclude_sample,exclude_sample],k = 10)
hc_cluster_ALL <- HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[exclude_sample,exclude_sample],k = 10,only_cluster = T)


colnames(hc_cluster_ALL) <- c("cluster","samples")


hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Han"]] <- "EAS"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "French"]] <- "WEur"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Bengali"]] <- "SA"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Tianyuan"]] <- "AncientEAS"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Satsurblia"]] <- "Satsurblia"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "AfontovaGora3"]] <- "ANE"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Yana1"]] <- "ANS"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Villabruna"]] <- "postLGM-WEurHG"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "SunghirI"]] <- "preLGM-WEurHG"
hc_cluster_ALL$cluster[hc_cluster_ALL$cluster == hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "Balochi"]] <- "SWA"
hc_cluster_ALL <-  inner_join(hc_cluster_ALL,hc_Supercluster_ALL,by="samples")

hc_cluster_ALL <- rbind(hc_cluster_ALL,Oceania_cluster,Not_clustered,Africa,AFfamily = data.frame(samples=c("AfanasievoSon2","AfanasievoFather","AfanasievoMother","AfanasievoSon1"),cluster= "Yamnaya",Supercluster= hc_cluster_ALL$Supercluster[hc_cluster_ALL$samples == "AfanasievoSon1"]))

HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[hc_cluster_ALL$samples[hc_cluster_ALL$cluster == "EAS"],hc_cluster_ALL$samples[hc_cluster_ALL$cluster == "EAS"]],
                                 k = 3)
hc_cluster_Asia <- HC_Dendrogram(res_f3_gtLH_Autosomes_matrix_red_dis[hc_cluster_ALL$samples[hc_cluster_ALL$cluster == "EAS"],hc_cluster_ALL$samples[hc_cluster_ALL$cluster == "EAS"]],
                                 k = 3,only_cluster = T)

colnames(hc_cluster_Asia) <- c("cluster","samples")

hc_cluster_ALL$cluster[hc_cluster_ALL$samples %in% hc_cluster_Asia$samples[hc_cluster_Asia$cluster == hc_cluster_Asia$cluster[hc_cluster_Asia$samples == "SpiritCave"]]] <- "Siberia&Americas"

hc_cluster_ALL$cluster[hc_cluster_ALL$samples == "AfanasievoSon1"] <- "Yamnaya"

hc_cluster_ALL <- hc_cluster_ALL %>% distinct()
```


## Plots of resulting clusters

### Dendogram

```{r eval=T,echo=F, result='hide'}

cluster_color_broad_all = data.frame(clusterF3D_broad_all = c("AncientEAS","ANE","ANS","EarlyOoA","EAS","North_Africa","Oceania","postLGM-WEurHG","preLGM-WEurHG","SA","Satsurblia","Siberia&Americas" ,"Sub-Sahara-Africa","SWA","WEur","Yamnaya"),
                                 cluster_color = 
                                   c("#009292FF","#FFCC00","#6DB6FFFF","#004949FF","#924900FF","#808080","#006DDBFF","#490092FF","#DB6D00FF","#920000FF","#FFB6DBFF","#24FF24FF","#000000FF","#B66DFFFF","#FF6DB6FF","#B6DBFFFF"))

Supercluster_color = data.frame(SuperclusterF3D = c("Africa","Americas&Asia","EarlyOoA","Oceania","WEurCAS"),Supercluster_color = c("#000000FF","#924900FF","#004949FF","#006DDBFF","#FF6DB6FF"))

pacman::p_load(tidyverse, ggdendro, RColorBrewer, plotly)

dendrogram <- as.dendrogram(hclust(as.dist(res_f3_gtLH_Autosomes_matrix_full_dis)))
dendrogram_data <- dendro_data(dendrogram)
dendrogram_segments <- dendrogram_data$segments 

dendrogram_ends <- dendrogram_segments %>%
  filter(yend == 0) %>% # filter for terminal dendrogram ends
  left_join(.,dendrogram_data$labels, by = "x") %>% # .$labels contains the row names from dist_matrix (i.e., sample_name)
  dplyr::rename(sample_name = label) %>% inner_join(.,hc_cluster_ALL,by=c("sample_name"="samples")) %>%
  inner_join(.,Supercluster_color,by=c("Supercluster"="SuperclusterF3D")) %>%
  inner_join(.,cluster_color_broad_all,by=c("cluster"="clusterF3D_broad_all"))


P_Dendrogram_Superclusters <- ggplot(segment(dendrogram_data)) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + 
  geom_segment(data = dendrogram_ends,
               aes(x=x, y=y.x, xend=xend, yend=yend, color = as.factor(Supercluster))) + 
  geom_text(aes(x = x, y = y, label = label, angle = 90, hjust = 1), data= label(dendrogram_data)) +
  scale_y_continuous(expand = c(0.3, 0)) +
  scale_color_manual(name = "Genetic Superclusters",labels = Supercluster_color$SuperclusterF3D, values = Supercluster_color$Supercluster_color) +
  theme_dendro() 
P_Dendrogram_Superclusters

P_Dendrogram_clusters <- ggplot(segment(dendrogram_data)) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + 
  geom_segment(data = dendrogram_ends,
               aes(x=x, y=y.x, xend=xend, yend=yend, color = as.factor(cluster))) + 
  geom_text(aes(x = x, y = y, label = label, angle = 90, hjust = 1), data= label(dendrogram_data)) +
  scale_y_continuous(expand = c(0.3, 0)) +
  scale_color_manual(name = "Genetic Clusters",labels = cluster_color_broad_all$clusterF3D_broad,values = cluster_color_broad_all$cluster_color) +
  theme_dendro() 
P_Dendrogram_clusters

ggsave(filename = paste(folder_path,"/Defining_genetic_populations_F3_dist_Dendrogram_clusters_using_1240k.png",sep=""),P_Dendrogram_clusters,device = "png",width = 14,height =  14)
ggsave(filename = paste(folder_path,"/Defining_genetic_populations_F3_dist_Dendrogram_Superclusters_using_1240k.png",sep=""),P_Dendrogram_Superclusters,device = "png",width = 14,height =  14)

```

### MDS

```{r eval=T,echo=F, result='hide'}
F3_MDS_frags <- as.data.frame(cmdscale(
  as.dist(res_f3_gtLH_Autosomes_matrix_full_dis),
  k=10,eig = T)$points)


F3_MDS_frags_anno <- F3_MDS_frags %>% mutate(sample_name=colnames(res_f3_gtLH_Autosomes_matrix_full)) %>%
  inner_join(.,Joint_Meta,by=c("sample_name"="F3_pop_annotation")) %>% filter(duplicated(sample_name) == FALSE) %>% 
  inner_join(.,hc_cluster_ALL,by=c("sample_name"="samples"))

write.csv(F3_MDS_frags_anno,paste0(folder_path,"/Genetic_clustering_F3_MDS_anno.csv"),quote = F,row.names = F)

MDS_Supercluster <- F3_MDS_frags_anno %>% filter(sample_name != "OaseNew") %>%
  ggplot(data= ., aes(x=V1, y=V2,color=Supercluster,label= sample_name,shape=Supercluster)) +
  geom_point() +
  xlab("K1") +
  ylab("K2") +
  scale_color_manual(name = "Genetic Superclusters",breaks = Supercluster_color$SuperclusterF3D,labels = Supercluster_color$SuperclusterF3D, values = Supercluster_color$Supercluster_color) +
  scale_shape_manual(name = "Genetic Superclusters",breaks = Supercluster_color$SuperclusterF3D,labels = Supercluster_color$SuperclusterF3D,values= c(1:length(unique(Supercluster_color$Supercluster_color))))+
  geom_text_repel(data= F3_MDS_frags_anno[F3_MDS_frags_anno$ML_BP_Mean > 0,], aes(label=sample_name),max.overlaps = 30,show.legend = FALSE) 

MDS_Supercluster

MDS_cluster <- F3_MDS_frags_anno  %>%
  ggplot(data= ., aes(x=V1, y=V2,color=cluster,shape=cluster)) +
  geom_point() +
  xlab("K1") +
  ylab("K2") +
  scale_color_manual(name = "Genetic Clusters",breaks = cluster_color_broad_all$clusterF3D_broad,labels = cluster_color_broad_all$clusterF3D_broad,values = cluster_color_broad_all$cluster_color) +
  scale_shape_manual(name = "Genetic Clusters",breaks = cluster_color_broad_all$clusterF3D_broad,labels = cluster_color_broad_all$clusterF3D_broad,values= c(1:length(unique(cluster_color_broad_all$cluster_color))))+
  geom_text_repel(data= F3_MDS_frags_anno[F3_MDS_frags_anno$ML_BP_Mean > 0,], aes(label=sample_name),max.overlaps = 30,show.legend = FALSE)
MDS_cluster

ggsave(filename = paste(folder_path,"/Defining_genetic_populations_F3_dist_MDS_clusters_using_1240k.png",sep=""),MDS_cluster,device = "png",width = 14,height =  14)
ggsave(filename = paste(folder_path,"/Defining_genetic_populations_F3_dist_MDS_Superclusters_using_1240k.png",sep=""),MDS_Supercluster,device = "png",width = 14,height =  14)

colnames(hc_cluster_ALL) <- c("clusterF3D_broad","samples","SuperclusterF3D")
Joint_Meta <- left_join(Joint_Meta,hc_cluster_ALL,by=c("F3_pop_annotation"="samples")) %>% distinct()
Joint_Meta$clusterF3D_broad[is.na(Joint_Meta$clusterF3D)] <- "Sub-Sahara-Africa"
Joint_Meta$SuperclusterF3D[is.na(Joint_Meta$SuperclusterF3D)] <- "Africa"

write.csv(Joint_Meta,paste(folder_path,"Joint_Meta_data_genetic_clusters_using_1240k.csv"),quote = F,row.names = F)

```


