---
  title: "Joint Admixture Dating Analysis"
output:
  html_document:
  keep_md: yes
pdf_document: default
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Joint Admixture Dating Analysis

## Functions

```{r eval=T,echo=F,results='hide'}
source("Analysis_R_functions.R")
```

## Meta Data

```{r eval=T,echo=F,results='hide'}
############## change path  ##################
folder_path = "Paper_figures"

folder_path_data = "Paper_figures_Data/"

folder_path_Sup_Tables = "Supplements_Tables/"

path_to_dryad_downloaded_files = "Dryad_folder"

path_to_my_admixfrog_folder_ancient = "Admixfrog_res_ancient"

path_to_my_admixfrog_folder_present = "Admixfrog_res_present"
##############################################

Joint_Meta <- read.csv(paste0(path_to_dryad_downloaded_files,"Meta_Data_individuals.csv")) %>% 
  filter(sample_name != "Mota",superpopulation_cluster != "Africa") %>% mutate(sample_name=gsub( "-", ".", sample_name))


cluster_color_broad_all = data.frame(clusterF3D_broad_all = c("AncientEAS","ANE","ANS","EarlyOoA","EAS","North_Africa","postLGM-WEurHG","preLGM-WEurHG","SA","Sahul","Satsurblia","Siberia&Americas" ,"Sub-Sahara-Africa","SWA","WEur","Yamnaya"),
                                     cluster_color = 
                                       c("#009292FF","#FFCC00","#6DB6FFFF","#004949FF","#924900FF","#808080","#490092FF","#DB6D00FF","#920000FF","#006DDBFF","#FFB6DBFF","#24FF24FF","#000000FF","#B66DFFFF","#FF6DB6FF","#B6DBFFFF"))

cluster_color_broad = data.frame(clusterF3D_broad = c("AncientEAS","ANE","ANS","EarlyOoA","EAS","postLGM-WEurHG","preLGM-WEurHG","SA","Sahul","Satsurblia","Siberia&Americas" ,"SWA","WEur","Yamnaya"),
                                 cluster_color = 
                                   c("#009292FF","#FFCC00","#6DB6FFFF","#004949FF","#924900FF","#490092FF","#DB6D00FF","#920000FF","#006DDBFF","#FFB6DBFF","#24FF24FF","#B66DFFFF","#FF6DB6FF","#B6DBFFFF"))

cluster_color_refined = data.frame(clusterF3D_refined = c("AncientEAS","ANE","ANS","EarlyOoA","EAS","EHG","IUP","preLGM-WEurHG","SA","Sahul","Satsurblia","Siberia&Americas" ,"SWA","WEur","WHG","Yamnaya"),
                                   cluster_color = c("#009292FF","#6DB6FFFF","#004949FF","#924900FF","#DB6D00FF","#808080","#920000FF","#006DDBFF","#FFB6DBFF","#24FF24FF","#000000FF","#B66DFFFF","#FF6DB6FF","#490092FF","#B6DBFFFF","#FFCC00"))


Supercluster_color = data.frame(SuperclusterF3D = c("Americas&Asia","EarlyOoA","Sahul","WEurCAS"),Supercluster_color = c("#924900FF","#004949FF","#006DDBFF","#FF6DB6FF"))

cluster_color = cluster_color_broad

gen_time = 28

# dates from Djakovic et al 2022 for Spain and France
Protoaurignacian_start = 42653
Chatelperronian_extinction = 39798
Neandertal_extinction_Djakovic = 40457

# Higham et al 2014
TransitionalIndustry_start = 43000
Nenadertal_extinction_Higham = 39220

Archeaology_first_humans = max(c(Protoaurignacian_start,TransitionalIndustry_start)) / gen_time
Archeaology_last_neandertals = min(c(Chatelperronian_extinction,Neandertal_extinction_Djakovic,Nenadertal_extinction_Higham)) / gen_time


```


## Simulations under the model

We used a snakemake script to do the simulation that calls 3 different R scripts. Parameters are read in from a yaml file. You can find the scripts in the Dating Neandertal ancestry folder on github in the Simulatiin folder.

```{r eval=T,echo=F, result='hide'}


path_SP <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "SP_sim_",full.names = T)

for(i in 1:length(path_SP)){
  load(path_SP[i])
  if(i == 1){
    SP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    SP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    SP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    SP_all_res = rbind(SP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    SP_res_matrix_all_Segfit <- rbind(SP_res_matrix_all_Segfit,res_matrix_Segfit)
    SP_res_matrix_all_ACov_fit_LL <- rbind(SP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }


}

SP_res_matrix_all_Segfit <- SP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

SP_res_matrix_all_ACov_fit_LL <- SP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))


SP_res_matrix_all <- rbind(SP_res_matrix_all_Segfit,SP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,SP_res_matrix_all_Segfit,SP_res_matrix_all_ACov_fit_LL)


path_TP <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "TP_sim_",full.names = T)

for(i in 1:length(path_TP)){
  load(path_TP[i])
  if(i == 1){
    TP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    TP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    TP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    TP_all_res = rbind(TP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    TP_res_matrix_all_Segfit <- rbind(TP_res_matrix_all_Segfit,res_matrix_Segfit)
    TP_res_matrix_all_ACov_fit_LL <- rbind(TP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

TP_res_matrix_all_Segfit <- TP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

TP_res_matrix_all_ACov_fit_LL <- TP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))


TP_res_matrix_all <- rbind(TP_res_matrix_all_Segfit,TP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,TP_res_matrix_all_Segfit,TP_res_matrix_all_ACov_fit_LL)


path_EP_1600_400 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1600_400",full.names = T)

for(i in 1:length(path_EP_1600_400)){
  load(path_EP_1600_400[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% dplyr::group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1600_400 = EP_all_res
EP_res_matrix_all_1600_400 <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)

path_EP_1600_1 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1600_1",full.names = T)

for(i in 1:length(path_EP_1600_1)){
  load(path_EP_1600_1[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1600_1= EP_all_res
EP_res_matrix_all_1600_1 <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)


path_EP_1600_200 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1600_200",full.names = T)

for(i in 1:length(path_EP_1600_200)){
  load(path_EP_1600_200[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1600_200 = EP_all_res
EP_res_matrix_all_1600_200 <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)

path_EP_1700_400 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1700_400",full.names = T)

for(i in 1:length(path_EP_1700_400)){
  load(path_EP_1700_400[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1700_400  = EP_all_res
EP_res_matrix_all_1700_400  <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)

path_EP_1800_600 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1800_600",full.names = T)

for(i in 1:length(path_EP_1800_600)){
  load(path_EP_1800_600[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1800_600 = EP_all_res
EP_res_matrix_all_1800_600  <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)


path_EP_1900_800 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_1900_800",full.names = T)

for(i in 1:length(path_EP_1900_800)){
  load(path_EP_1900_800[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_1900_800 = EP_all_res
EP_res_matrix_all_1900_800  <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)

path_EP_2000_1000 <- list.files(path = "EMH_dating_analysis/Test_Sim/",pattern = "EP_sim_0.015_tm_td_comb_2000_1000",full.names = T)

for(i in 1:length(path_EP_2000_1000)){
  load(path_EP_2000_1000[i])
  if(i == 1){
    EP_all_res = rbind(Sp_ACov_res,Sp_seg_res)
    EP_res_matrix_all_ACov_fit_LL = res_matrix_ACov_fit_LL
    EP_res_matrix_all_Segfit = res_matrix_Segfit
  } else {
    EP_all_res = rbind(EP_all_res,rbind(Sp_ACov_res,Sp_seg_res))
    EP_res_matrix_all_Segfit <- rbind(EP_res_matrix_all_Segfit,res_matrix_Segfit)
    EP_res_matrix_all_ACov_fit_LL <- rbind(EP_res_matrix_all_ACov_fit_LL,res_matrix_ACov_fit_LL)
  }

}

EP_res_matrix_all_Segfit <- EP_res_matrix_all_Segfit %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_res_matrix_all_ACov_fit_LL <- EP_res_matrix_all_ACov_fit_LL %>% group_by(Var1,Var2,method,model,simulation,true_tm1,true_tm2,true_td,truncation,noise) %>%
  dplyr::summarize(value = sum(value), value0 = mean(value0)) %>% ungroup() %>% group_by(noise) %>%
  mutate(est_td = Var1[which.max(value0)],est_tm = Var2[which.max(value0)]) %>%
  mutate(value0 = ifelse(value0 < -100,-101,value0))

EP_all_res_2000_1000 = EP_all_res
EP_res_matrix_all_2000_1000  <- rbind(EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)
rm(Sp_ACov_res,Sp_seg_res,res_matrix_Segfit,res_matrix_ACov_fit_LL,EP_res_matrix_all_Segfit,EP_res_matrix_all_ACov_fit_LL)


P_SP_fit_res_A_Data <- rbind(SP_all_res,TP_all_res,EP_all_res_1600_400) %>% 
  mutate(sim = ifelse(simulation == "SimplePulse","Simple Pulse",ifelse(simulation == "TwoPulses","Two Pulse","Extended Pulse"))) 

P_SP_fit_res_B_Data <- rbind(EP_all_res_1600_1,EP_all_res_1600_200,EP_all_res_1700_400,EP_all_res_1800_600,EP_all_res_1900_800,EP_all_res_2000_1000) %>% 
  mutate(sim = ifelse(simulation == "SimplePulse","Simple Pulse",ifelse(simulation == "TwoPulses","Two Pulse","Extended Pulse"))) 

P_EP_fit_res_C_Data <- rbind(SP_res_matrix_all,TP_res_matrix_all,EP_res_matrix_all_1600_400) %>% 
  mutate(sim = ifelse(simulation == "SimplePulse","Simple Pulse",ifelse(simulation == "TwoPulses","Two Pulse","Extended Pulse"))) 

P_EP_fit_res_D_Data <- rbind(EP_res_matrix_all_1600_1,EP_res_matrix_all_1600_200,EP_res_matrix_all_1700_400,EP_res_matrix_all_1800_600,EP_res_matrix_all_1900_800,EP_res_matrix_all_2000_1000) %>% 
  mutate(sim = ifelse(simulation == "SimplePulse","Simple Pulse",ifelse(simulation == "TwoPulses","Two Pulse","Extended Pulse")))

write.csv(P_SP_fit_res_A_Data,paste0(folder_path_data,"P_SP_fit_res_A_Data.csv"),quote = F,row.names = F)
write.csv(P_SP_fit_res_B_Data,paste0(folder_path_data,"P_SP_fit_res_B_Data.csv"),quote = F,row.names = F)
write.csv(P_EP_fit_res_C_Data,paste0(folder_path_data,"P_EP_fit_res_C_Data.csv"),quote = F,row.names = F)
write.csv(P_EP_fit_res_D_Data,paste0(folder_path_data,"P_EP_fit_res_D_Data.csv"),quote = F,row.names = F)

```

## Estimating Jackknife se by chrom for SP

All input files here are obtained by running the Dating_Snakemake_pipeline.py script.

```{r eval=T,echo=F, result='hide'}


## ACov

JK_ACov <- read.csv("Dating_Results/ACov_Datinf_asc.snp",sep=" ",header = F) %>% 
  dplyr::select(V1,V2) %>% group_by(V1) %>% dplyr::summarize(n_pos = n())


SP_Acov <- read.csv("Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_ACov_SP_res_merged_ACov_Run_SP.csv") %>% mutate( JK = ifelse(JKchrom == "All", F,T), exclude = ifelse(exclude == "", "All_Data", exclude)) 

ACov_Run_SP_JKchrom_mean_out <- list()
est = unique(SP_Acov[,c("truncation","weighted","exclude")])
for(i in 1:length(est$truncation)){
  ACov_Run_SP_JKchrom_summary <- SP_Acov %>% 
    filter(truncation == est$truncation[i], weighted == est$weighted[i], exclude == est$exclude[i], JK == T) %>% 
    mutate(JKchrom = as.numeric(JKchrom)) %>%
    inner_join(.,JK_ACov,by=c("JKchrom"="V1"))
  ACov_Run_SP_JKchrom_mean <- SP_Acov %>% filter(truncation == est$truncation[i], weighted == est$weighted[i], exclude == est$exclude[i], JK == F) 
  
  write.table(x = data.frame(ACov_Run_SP_JKchrom_summary$JKchrom,ACov_Run_SP_JKchrom_summary$n_pos,ACov_Run_SP_JKchrom_summary$tm),file = paste0("Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/ACov_Run_EP_JKchrom_sum_tm_",est$truncation[i],"_",est$weighted[i],"_",est$exclude[i],".jin"),quote = F,row.names = F,col.names = F)

  system(paste0("~/DATES/src/./dowtjack -i Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/ACov_Run_EP_JKchrom_sum_tm_",est$truncation[i],"_",est$weighted[i],"_",est$exclude[i],".jin -m ",ACov_Run_SP_JKchrom_mean$tm," -o Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/ACov_Run_EP_JKchrom_sum_tm_",est$truncation[i],"_",est$weighted[i],"_",est$exclude[i],".jout"))

  tm_SE_est_SP <- read.table(paste0("Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/ACov_Run_EP_JKchrom_sum_tm_",est$truncation[i],"_",est$weighted[i],"_",est$exclude[i],".jout"))
  ACov_Run_SP_JKchrom_mean_out[[i]] <- ACov_Run_SP_JKchrom_mean %>% mutate(mean_JK = tm_SE_est_SP[1,1],se_JK = tm_SE_est_SP[1,2])

}

SP_Acov_anno <- do.call("rbind",ACov_Run_SP_JKchrom_mean_out) %>%
  mutate(weighted2 = ifelse(weighted == "yes","Same weight for each Individual","Weight proportional to data")) %>%
  dplyr::select(tm,mean_JK,se_JK,tm_lower,tm_upper,LL,map,truncation,weighted2,exclude,gen_time) %>% 
  dplyr::rename(weighted = weighted2) %>% 
  filter(map == "Shared_Map")

write.table(SP_Acov_anno,"Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_ACov_SP_res_merged_ACov_Run_SP_annotated.csv",quote = F,row.names = F,sep=",")

## Seg

JK_segments <- read.csv(paste0(path_to_dryad_downloaded_files,"/ref_archaicadmixtureAPX_hs37mMask35to99.csv.xz")) %>%
  filter(chrom != "X") %>%
  group_by(chrom) %>% dplyr::summarize(chrom_start = min(pos),chrom_end = max(pos)) %>% mutate(chrom_length = chrom_end - chrom_start)

SP_Seg <- read.csv("Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_Seg_SP_res_merged_Seg_Run_SP.csv") %>% 
  mutate( JK = ifelse(JKchrom == "All", F,T), exclude = ifelse(exclude == "", "All_Data", exclude)) 

Seg_Run_SP_JKchrom_mean_out <- list()
est = unique(SP_Seg[,c("min_l","exclude")])
for(i in 1:length(est$min_l)){
  Seg_Run_SP_JKchrom_summary <- SP_Seg %>% 
    filter(exclude == est$exclude[i], min_l == est$min_l[i],JK == T) %>% 
    inner_join(.,JK_segments,by=c("JKchrom"="chrom"))
  Seg_Run_SP_JKchrom_mean <- SP_Seg %>% filter(exclude == est$exclude[i], min_l == est$min_l[i],JK == F)
  
  write.table(x = data.frame(Seg_Run_SP_JKchrom_summary$JKchrom,Seg_Run_SP_JKchrom_summary$chrom_length,Seg_Run_SP_JKchrom_summary$tm),file = paste0("Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/Seg_Run_EP_JKchrom_sum_tm_",est$min_l[i],"_",est$exclude[i],".jin"),quote = F,row.names = F,col.names = F)

  system(paste0("~/DATES/src/./dowtjack -i Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/Seg_Run_EP_JKchrom_sum_tm_",est$min_l[i],"_",est$exclude[i],".jin -m ",Seg_Run_SP_JKchrom_mean$tm," -o Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/Seg_Run_EP_JKchrom_sum_tm_",est$min_l[i],"_",est$exclude[i],".jout"))

  tm_SE_est_SP <- read.table(paste0("Dating_Results/Fit_data_tm_step_1_td_step_1/JKResults/Seg_Run_EP_JKchrom_sum_tm_",est$min_l[i],"_",est$exclude[i],".jout"))
  Seg_Run_SP_JKchrom_mean_out[[i]] <- Seg_Run_SP_JKchrom_mean %>% mutate(mean_JK = tm_SE_est_SP[1,1],se_JK = tm_SE_est_SP[1,2])

}

SP_Seg_anno <- do.call("rbind",Seg_Run_SP_JKchrom_mean_out) %>%
  mutate(LL = -LL,min_l = as.factor(min_l)) %>%
  dplyr::select(tm,mean_JK,se_JK,tm_lower,tm_upper,LL,map,min_l,exclude)  %>% filter(map == "Shared_Map")

write.csv(SP_Seg_anno,"Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_Seg_SP_res_merged_Seg_Run_SP_annotated.csv",quote = F,row.names = F)
```



Estimating it from the range of LL being bigger than -2


```{r eval=T,echo=F, result='hide'}
ACov_Run_EP  <- read.csv("Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_ACov_EP_res_merged_ACov_Run_EP.csv") %>% 
  mutate(exclude2 = ifelse(exclude == "", "All Data",ifelse(startsWith(exclude,"BKBB7240_BKF6620"), "EarlyOoA Excluded",ifelse(exclude == "OaseNew","Oase Excluded",ifelse(startsWith(exclude,"OaseNew_PM1"),"Only EarlyOoA w/o Oase","Only EarlyOoA"))))) 
  group_by(truncation,exclude,min_l,map,weighted) %>% filter(value0 > -101) 

write.csv(x=ACov_Run_EP,file = "Dating_Results/Fit_data_tm_step_1_td_step_1/Results/All_ACov_EP_res_merged_ACov_Run_EP_reduced.csv",quote = F,row.names = F)

ACov_Run_EP_summary <- ACov_Run_EP %>% group_by(truncation,exclude,min_l,map,weighted) %>% filter(value0 == 0) %>% ungroup()

ACov_Run_EP_conditioned <- inner_join(ACov_Run_EP,ACov_Run_EP_summary[,c("truncation","exclude2","weighted","Var2")],
                                      by=c("truncation"="truncation","exclude2"="exclude2","weighted"="weighted","Var2"="Var2")) 

ACov_Run_EP_conditioned_summary <- ACov_Run_EP_conditioned %>% filter(value0 > -2) %>%
  group_by(truncation,exclude,min_l,map,weighted) %>% 
  dplyr::summarize(tm_mean = mean(Var2),tm_min = min(Var2),tm_max = max(Var2),td_mean = mean(Var1),td_min = min(Var1),td_max = max(Var1)) %>% ungroup() %>%
  dplyr::select(truncation,exclude,min_l,map,weighted,tm_mean,tm_min,tm_max,td_mean,td_min,td_max) %>%
  inner_join(ACov_Run_EP_summary,.) %>% dplyr::select(!exclude) %>% dplyr:: rename(exclude = exclude2)

ACov_Run_EP_summary <- ACov_Run_EP %>% filter(value0 > -2) %>% group_by(truncation,exclude,min_l,map,weighted) %>% 
  dplyr::summarize(tm_mean = mean(Var2),tm_min = min(Var2),tm_max = max(Var2),td_mean = mean(Var1),td_min = min(Var1),td_max = max(Var1)) %>% ungroup() %>%
  dplyr::select(truncation,exclude,min_l,map,weighted,tm_mean,tm_min,tm_max,td_mean,td_min,td_max) %>%
  inner_join(ACov_Run_EP_summary,.) %>% dplyr::select(!exclude) %>% dplyr:: rename(exclude = exclude2)

P_LL_td_conditioned_tm <- ACov_Run_EP_conditioned %>% 
  mutate(weighted = ifelse(weighted == "yes","weighted equally","unweighted")) %>%
  ggplot(.,aes(x=Var1,y=value0,color = truncation)) +
    geom_line() +
    facet_wrap(exclude2~weighted,scales = "free",ncol = 2) +
    geom_hline(yintercept = -2) +
    coord_cartesian(expand = F,ylim = c(-20,2))



```



